---
title: "hw2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(gapminder)
library(reshape2)
library(maps)
library(ggmap)
```

## Loading and modifying data

```{r}
data <- read.csv(file = './data/fec_2008-2022.csv')

data <- data %>% 
  select(-c(X))

# Column with only numerical and int data type
num_data <- data %>% 
  select(where(is.numeric)) %>% 
  select(-c(Cand_Office_Dist, Cand_Zip))

# Convert party affiliation to number -1 to 1 to use later for visualization
data$Cand_State <- tolower(state.name[match(data$Cand_State, state.abb)])
data$Cand_Office_St <- tolower(state.name[match(data$Cand_Office_St, state.abb)])

data <- data %>% mutate(
  Cand_Party_Affiliation_num = case_when(Cand_Party_Affiliation=="DEM" ~ 1,
                                            Cand_Party_Affiliation=="REP" ~ -1,
            Cand_Party_Affiliation!="REP" | Cand_Party_Affiliation!="DEM" ~ 0))


# Check data
# summary(data)
str(data)

```


## Correlation


```{r}

num_data$Debt_Owe_To_Committee[is.na(num_data$Debt_Owe_To_Committee)] <- median(num_data$Debt_Owe_To_Committee, na.rm=TRUE)

num_data$Debt_Owed_By_Committee[is.na(num_data$Debt_Owed_By_Committee)] <- median(num_data$Debt_Owed_By_Committee, na.rm=TRUE)

num_data$Cash_On_Hand_BOP[is.na(num_data$Cash_On_Hand_BOP)] <- median(num_data$Cash_On_Hand_BOP, na.rm=TRUE)

num_data$Cash_On_Hand_COP[is.na(num_data$Cash_On_Hand_COP)] <- median(num_data$Cash_On_Hand_COP, na.rm=TRUE)


# Get correlation matrix
corr <- cor(num_data)

# Fill upper triangle to NaN values
corr[upper.tri(corr)] <- NA

# Melt to one to one dataframe
corr <- melt(corr)

# corr

# Part for sorting unused
# group_corr <- corr %>% 
#   drop_na() %>% 
#   group_by(Var2) %>% 
#   summarize(med_value= median(value))
# 
# group_corr
# 
# corr <- corr %>% 
#   left_join(group_corr, by="Var2")
# 
# corr <- corr[order(corr$med_val, corr$Var2),]

```



```{r, fig1, fig.height = 10, fig.width = 12}
ggplot(data=corr, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile(color="white") +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_gradient2(low="blue", mid="white", high="red", na.value = "grey90",
                       limit = c(-1,1), name="Pearson Correlation")
```

Type of information:
1. correlation between variables.

Findings:
1. Total receipt, total disbursement, total contribution, and operating expenditure have very high correlation.
2. All variables have higher than 0 correlation each other.


## Candidate VS Total_Receipt

```{r, fig2, fig.height = 13, fig.width = 12}

# Sort by Total_Receipt and get top 50 candidates
receipt_data = data[order(-data$Total_Receipt),]
receipt_data = receipt_data[0:50,]

ggplot(receipt_data, aes(x=reorder(Cand_Name,Total_Receipt), y=Total_Receipt)) +
  geom_point() +
  coord_flip() +
  theme(text = element_text(size=13, color="black"), axis.text.y=element_text(colour="black"))
```

Type of information:
1. Trends of total receipt
2. Who get more receipt

Findings
1. A few famous people dominates total receipt.(exponentially increases)
2. Large variance for top ten people, small variance for others.


## Party Affiliation Ratio by states

```{r}
# Get party_ratio on each states
party_ratio_states <- data %>%
  group_by(Cand_Office_St) %>%
  summarize(party_ratio = mean(Cand_Party_Affiliation_num)) %>% 
  select(c(Cand_Office_St, party_ratio))

head(party_ratio_states)
```


```{r}

# Get map of united states consisting of longitude and latitude
states_map <- map_data("state")

head(states_map)

map_data <- left_join(party_ratio_states , states_map, by=c("Cand_Office_St" = "region"))

head(map_data)

```


```{r, fig2.5, fig.height = 10, fig.width = 17}
ggplot(map_data, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = party_ratio), color = "white") +
    scale_fill_gradient2(limit = c(-1,1),low="blue", mid="white", high="red", 
                         na.value = "grey90", name="Party Ratio")
```


Type of information:
1. Ratio of office for party affiliation in each state


What we can see:
1. Most regions are close to 0, so that means number of offices for each party in each state is not that biased.
2. A few democratic
3. More Democratic offices on West and East coast.


## Party Affiliations and their numbers


```{r, fig3, fig.height = 17, fig.width = 10}
data2 <- data %>% 
  group_by(Cand_Party_Affiliation) %>% 
  mutate(count=n())

ggplot(data2, aes(x=reorder(Cand_Party_Affiliation, count), y=count)) +
  geom_bar(stat="identity", fill="skyblue") +
  coord_flip()
```

Type of information:
1. Number of party affiliations in US.
2. Trends of support for party affiliations

What we can see:

1. There are so many party affiliations other than Republic and Democratic with almost 0 people.
2. Majorities are either Republcan, Democratic, or Independent.
3. There are so many party affiliations other than Republic and Democratic.














