---
title: "hw2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(gapminder)
library(reshape2)
library(maps)
library(ggmap)
library(lubridate)
```

## Loading and modifying data

```{r}
data <- read.csv(file = './data/fec_2008-2022.csv')

# Get rid of variable X from data
data <- data %>% 
  select(-X)

# Column with only numerical and int data type
num_data <- data %>% 
  select(where(is.numeric)) %>% 
  select(-c(Cand_Office_Dist, Cand_Zip))

# Convert party affiliation to number -1 to 1 to use later for visualization
data$Cand_State <- tolower(state.name[match(data$Cand_State, state.abb)])
data$Cand_Office_St <- tolower(state.name[match(data$Cand_Office_St, state.abb)])

data <- data %>% mutate(
  Cand_Party_Affiliation_num = case_when(Cand_Party_Affiliation=="DEM" ~ 1,
                                            Cand_Party_Affiliation=="REP" ~ -1,
            Cand_Party_Affiliation!="REP" | Cand_Party_Affiliation!="DEM" ~ 0))


# Check data
# summary(data)
str(data)

```


## Correlation

```{r}

num_data$Debt_Owe_To_Committee[is.na(num_data$Debt_Owe_To_Committee)] <- median(num_data$Debt_Owe_To_Committee, na.rm=TRUE)

num_data$Debt_Owed_By_Committee[is.na(num_data$Debt_Owed_By_Committee)] <- median(num_data$Debt_Owed_By_Committee, na.rm=TRUE)

num_data$Cash_On_Hand_BOP[is.na(num_data$Cash_On_Hand_BOP)] <- median(num_data$Cash_On_Hand_BOP, na.rm=TRUE)

num_data$Cash_On_Hand_COP[is.na(num_data$Cash_On_Hand_COP)] <- median(num_data$Cash_On_Hand_COP, na.rm=TRUE)


# Get correlation matrix
corr <- cor(num_data)


# Fill upper triangle to NaN values
corr[lower.tri(corr)] <- NA

# Melt to one to one dataframe
corr <- melt(corr)
```



```{r, fig1, fig.height = 10, fig.width = 12}
ggplot(data=corr, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile(color="white") +
  scale_x_discrete(position = "top") +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle = 90, hjust=0, vjust=0.5)) +
  scale_fill_gradient2(low="blue", mid="white", high="red", na.value = "grey90",
                       limit = c(-1,1), name="Pearson Correlation")
```

Type of information:
1. correlation between variables.

Findings:
1. Total receipt, total disbursement, total contribution, and operating expenditure have very high correlation.
2. All variables have higher than 0 correlation each other.


## Candidate VS Total_Receipt

```{r, fig2, fig.height = 8, fig.width = 8}

# Sort by Total_Receipt and get top 50 candidates
receipt_data = data[order(-data$Total_Receipt),]
receipt_data = receipt_data[0:40,]

ggplot(receipt_data, aes(x=reorder(Cand_Name,Total_Receipt), y=Total_Receipt/1000000)) +
  geom_point() +
  coord_flip() +
  theme(text = element_text(size=11, color="black"), axis.text.y=element_text(colour="black"),
        axis.title=element_text(size=12)) +
  labs(x="Candidate Name", y="Total Receipt in million dollars")

```

Type of information:
1. Trends of total receipt
2. Who get more receipt

Findings
1. A few famous people dominates total receipt.(exponentially increases)
2. Large variance for top ten people, small variance for others.


## Party Affiliation Ratio by states

```{r}
# Get party_ratio on each states
party_ratio_states <- data %>%
  group_by(Cand_Office_St) %>%
  summarize(party_ratio = mean(Cand_Party_Affiliation_num)) %>% 
  select(c(Cand_Office_St, party_ratio))

head(party_ratio_states)
```


```{r}

# Get map of united states consisting of longitude and latitude
states_map <- map_data("state")

head(states_map)

map_data <- left_join(party_ratio_states , states_map, by=c("Cand_Office_St" = "region"))

head(map_data)

```


```{r, fig2.5, fig.height = 6.2, fig.width = 11}
ggplot(map_data, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = party_ratio), color = "white") +
    scale_fill_gradient2(limit = c(-1,1),low="blue", mid="white", high="red", 
                         na.value = "grey90", name="Party Ratio") + 
  theme(panel.background = element_rect(fill = 'grey70', colour = 'white'))
```


Type of information:
1. Ratio of office for party affiliation in each state


What we can see:
1. Most regions are close to 0, so that means number of offices for each party in each state is not that biased.
2. A few democratic
3. More Democratic offices on West and East coast.


## Party Affiliations and their numbers


```{r, fig3, fig.height = 8, fig.width = 5}
data2 <- data %>% 
  group_by(Cand_Party_Affiliation) %>% 
  summarize(count=n())

data2 <- data2 %>% drop_na() %>% 
  mutate(Cand_Party_Affiliation = ifelse(count <= 30, "Others (113 parties)", data2$Cand_Party_Affiliation))


ggplot(data2, aes(x=reorder(Cand_Party_Affiliation, count), y=count)) +
  geom_bar(stat="identity", fill="skyblue") +
  coord_flip() +
  labs(x="") +
  ggtitle("Number of candidates in each party") +
  theme(text=element_text(size=9), plot.title = element_text(hjust = 0.5))
```

Type of information:
1. Number of party affiliations in US.
2. Trends of support for party affiliations

What we can see:

1. There are so many party affiliations other than Republic and Democratic with almost 0 people.
2. Majorities are either Republcan, Democratic, or Independent.
3. There are so many party affiliations other than Republic and Democratic.


## Part 2

Hypothesis: Interest toward US politics increased significantly in 2020 election.

From the plot of total receipt divided by winner and loser in each year, we can see that the total amount of receipt was the highest in 2020. This may mean that more people donated their money to candidates in 2020. We will investigate this by seeing amount of individual contributions in each year.

```{r}
# Extract Year from date
data$End_Year = year(as.Date(data$Coverage_End_Date, format = "%m/%d/%Y"))
```

```{r}
year_cont <- data %>% 
  select(End_Year, Individual_Contribution)

unique(year_cont$End_Year)

year_cont <- year_cont %>% mutate(
  End_Year= case_when(End_Year >= 2009 & End_Year <= 2012 ~ 2012,
                      End_Year >= 2013 & End_Year <= 2016 ~ 2016,
                      End_Year >= 2017 & End_Year <= 2020 ~ 2020))



year_cont <- year_cont %>% group_by(End_Year) %>% 
  drop_na() %>% 
  filter(End_Year > 2008 & End_Year < 2021) %>% 
  summarize(Sum_Contribution = sum(Individual_Contribution)/1000000)

year_cont <- as.data.frame(year_cont)

head(year_cont)
```



```{r, fig6, fig.height = 4, fig.width = 4.5}
ggplot(data=year_cont, aes(x=End_Year, y=Sum_Contribution)) +
  geom_bar(stat="identity", fill="steelblue") + 
  scale_x_continuous(breaks= c(2012, 2016, 2020)) +
  labs(x="Year", y="") +
  ggtitle("Individual Contribution in One million dollars") + 
  theme(text=element_text(size=9), plot.title = element_text(hjust = 0.5))
  # scale_fill_discrete(labels="Individual Contribution") +
  # theme(legend.position=c(0.3, 0.8), legend.background=element_rect(fill="transparent"))
  
```

Both 2012 and 2016 doesn't look that different, but individual contribution increased almost two times as much as those in 2012 and 2016.This shows that public interest toward politics went much higher because many people donated more money to candidates.









